---
title: "Longevity after Elections"
author: "Lela Sengupta"
format: html
execute:
  echo: false
---

```{r}
#|label: setup
#|message: false
#|warning: false
library(tidyverse)
library(primer.data)
library(brms)
library(tidybayes)
library(gtsummary)
```

```{r}
#|label: data_load
#|message: false

x <- governors |> 
  #selecting 4 columns that connect with the model
  select(death_age, election_age, win_margin, party, sex) |> 
  #make new column to say win if win margin > 0, lose if not
  mutate(treatment = ifelse(win_margin > 0, "win", "lose")) |> 
  #keeping candidates who had a win margin of less than 5
  filter(abs(win_margin) < 5) 
 

```

```{r}
#|label: model
#|cache: true

fit_elec <- brm(formula = death_age ~ treatment + party + win_margin,
     data = x,
     family = gaussian(),
     refresh = 0,
     silent = 2,
     seed = 23
     )
```

$$death\_age_{i} = \beta_{0} + \beta_{1} treatment_i + \beta_{2}win\_margin_i + \beta_{3}party_i + \epsilon_{i}$$

```{r}
#| warning: false
tbl_regression(fit_elec)
```



## Wisdom

subject: longevity

broad q: election results and age

specific q: how long will preceptor live if he wins the election for mayor of newton? 

timeframe: in the near future (2024-2050)

preceptor table:
- units: candidates for mayor
- potential outcomes: number of years alive after election,
number of years alive after losing election (lived_after)
- covariates: sex (sex), age when elected (election_age)
- treatment: won or lost election (win_margin)

model type: gaussian

validity:
- winning a municipal election is different from winning a state election

## Justice
population table:
- years: 1945-2050 (when candidates run for election)
- units: any candidate running for a political position in the US
- potential outcomes: number of years alive after election,
number of years alive after losing election (lived_after)
- covariates: sex (sex), age when elected (election_age)
- treatment: won or lost election (win_margin)

stability:
- life expectancy has changed since our data was collected

representativeness:
- we want data for mayoral elections, which are different from state elections
- we have less data for losing candidates than winning candidates

unconfoundedness:
- many covariates influence whether a candidate wins or loses an election, including those that cannot be manipulated (race, sex, and age are some examples) 

##Courage
- people who have larger win margins are more likely to have other variables confound their results

models
testing
dgm


